---
title: "Reproducible Research Project 1"
author: "Giresse"
date: '2022-08-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(magrittr)
library(ggplot2)
```

## Loading and processing the data


```{r}
# download file from web
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "activity.zip",mode="wb")
# unzip data and read 
unzip("activity.zip")
mydf <- read.csv("activity.csv", header = TRUE)
head(mydf)
```

## What is mean total numbers of step taken per day

```{r echo=FALSE}
# Calculate the total steps taken per day:
total_by_day <- mydf %>% select(date, steps) %>% group_by(date) %>% summarize(total_steps= sum(steps)) %>%na.omit()

# A histogram of total steps taken per day:
hist(total_by_day$total_steps, xlab = "Total daily Steps",main="Histogram of Total Steps by day", breaks = 20)
```

## Calculate and report the mean and median of the total number of steps taken per day

```{r}
#mean of the total numbers of steps taken per day
mean(total_by_day$total_steps)

#mediane of the total numbers of steps taken per day
median(total_by_day$total_steps)
```

## Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
#Interval data

interval_data <- mydf %>% select(interval, steps) %>% na.omit() %>% group_by(interval) %>% summarize(min_interval= mean(steps)) 
  
#Time series plot of the 5-minute interval
ggplot(interval_data, aes(x=interval, y=min_interval))+ geom_line()

```

## The 5-minute interval that, on average, contains the maximum number of steps
```{r}
interval_data[which(interval_data$min_interval == max(interval_data$min_interval)),]
```

##Imputing missing values

```{r}
#Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with \color{red}{\verb|NA|}NAs)
missing_Values <- sum(is.na(mydf))
missing_Values

```

## Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r}
# Using the mean for that 5 -minute interval to replace all the missing values in the dataset

mean_replace_function <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
step_mean <- mydf %>% group_by(interval) %>% mutate(steps = mean_replace_function(steps))
sum(is.na(step_mean))

```

## Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
#Creating dataframe of steps by day

step_by_day <- aggregate(step_mean$steps, by=list(step_mean$date), sum)

step_by_day <- rename(step_by_day, Date = Group.1, Total_steps = x)

head(step_by_day)

# Calculating the mean and median of number steps taken per day

summary(step_by_day)

# Making histogram

hist(step_by_day$Total_steps,  xlab = "Steps", ylab = "Frequency", main = "Total Daily Steps", breaks = 20)


#Comparing the mean of total_by_day and step_by_day

mean_by_total_day = mean(total_by_day$total_steps, na.rm = TRUE)
mean_by_total_day
mean_step_by_day = mean(step_by_day$Total_steps)
mean_step_by_day 

#Comparing the median of total_by_day and step_by_day

median_by_total_day = median(total_by_day$total_steps, na.rm = TRUE)
median_by_total_day
median_step_by_day = median(step_by_day$Total_steps)
median_step_by_day 

```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
step_mean$date <- as.Date(step_mean$date)
step_mean$weekday <- weekdays(step_mean$date)
step_mean$weekend <- ifelse(step_mean$weekday == "Saturday" | step_mean$weekday == "Sunday", "Weekend", "Weekday")



```












